Characterising Optimisations to Memory Access Patterns using Architecture Independent Program Features

### 2020-01-16 paper/paper.tex
propose new abstract

	\begin{abstract}
	This paper addresses the challenge of profiling a code base to help programmers effectively optimise their codes for high performance.
	Current methods of analysing program behaviour have largely relied on performance counters on various architectures to provide architecture dependent metrics of code behaviour. 
	However, if workload characteristics are tied to architectural features specific to a particular system, these analysis frameworks may not generalise to alternative or future systems.
	Considering the trend towards heterogeneous computing, we work towards a framework to guide developers in optimising codes for any given hardware target using the architecture independent workload characterisation (AIWC) tool -- specifically in profiling memory access patterns of code. 
	Metrics must be easily interpreted and conceptually linked to patterns in hardware-specific memory access optimisations.
	To this end we propose, implement and analyse the addition of two new metrics to AIWC and present a new approach to profile spatial locality intrinsic to parallel codes.
	The improvements suggested to AIWC will help HPC developers better understand memory access patterns of complex codes, a crucial step in guiding optimisation of codes for arbitrary hardware targets.
	\end{abstract}
	\maketitle

@@

	High-performance computing (HPC) systems are increasingly incorporating a multitude of heterogeneous compute elements. Single nodes on supercomputers now comprise of traditional CPUs and multiple accelerators such as GPUs, Field Programmable Gate Arrays (FPGAs), and many-integrated-core devices (MICs). The development of high bandwidth interconnects between compute devices supports tight integration between multiple devices of different types on the same compute node.

@@

	Application codes differ in resource requirements, control structure and available parallelism. Similarly, compute devices differ in number and capabilities of execution units, processing model, and available resources. These heterogeneous computing environments present opportunities for software engineers and HPC integrators to design highly optimised systems with multiple kernels executing on hardware targets best suited for the diverse computational tasks performed by these kernels \cite{spafford2010maestro}.	However this opportunity raises the twin challenges of deciding on the right hardware target for complex kernels; and optimising codes to run on these diverse architectures. Our work is directed at helping developers tackle the latter of these two problems.

>>>>

	\begin{abstract}
	proposed new abstract:

	Context: AIWC gathers features of OpenCL programs which can be used to understand and predict program performance

	Problem: metrics aren't easily interpreted and don't reflect important memory access patterns that affect efficiency across all architectures

	Solution: new metrics

	Results: differences in metrics reflect optimizations for matrix multiply; improved performance prediction by x\% for Extended OpenDwarfs benchmarks

	Significance: as per last sentence\\

	This paper addresses the challenge of profiling a code base to help programmers effectively optimise their codes for high performance.
	Current methods of analysing program behaviour have largely relied on performance counters on various architectures to provide architecture dependent metrics of code behaviour. 
	However, if workload characteristics are tied to architectural features specific to a particular system, these analysis frameworks may not generalise to alternative or future systems.
	Considering the trend towards heterogeneous computing, we work towards a framework to guide developers in optimising codes for any given hardware target using the architecture independent workload characterisation (AIWC) tool -- specifically in profiling memory access patterns of code. 
	Metrics must be easily interpreted and conceptually linked to patterns in hardware-specific memory access optimisations.
	To this end we propose, implement and analyse the addition of two new metrics to AIWC and present a new approach to profile spatial locality intrinsic to parallel codes.
	The improvements suggested to AIWC will help HPC developers better understand memory access patterns of complex codes, a crucial step in guiding optimisation of codes for arbitrary hardware targets.
	\end{abstract}
	\maketitle

@@

	High-performance computing (HPC) systems are increasingly heterogeneous.
	A single node on a modern supercomputers may combine a traditional CPUs with one or more accelerators such as GPUs, Field Programmable Gate Arrays (FPGAs), or many integrated core devices (MICs).
	The development of high bandwidth interconnects between compute devices supports tight integration between multiple devices of different types on the same compute node.

@@

	Application codes differ in resource requirements, control structure and available parallelism. Similarly, compute devices differ in number and capabilities of execution units, processing model, and available resources. These heterogeneous computing environments present opportunities for software engineers and HPC integrators to design highly optimised systems with multiple kernels executing on hardware targets best suited for the diverse computational tasks performed by these kernels \cite{spafford2010maestro}.
	However this opportunity raises the twin challenges of determining the best hardware target for complex kernels and optimising codes to run on these diverse architectures.
	Our work is directed at helping developers tackle the latter of these two problems.
